---
title: "Denisty plot"
parent: Presentation
has_children: false
nav_order: 1
mathjax: true ## Switch to false if this page has no equations or other math rendering.
---

## Introduction

A Density Plot visualises the distribution of data over a continuous interval (or time period). Density Plots are not affected by the number of bins (each bar used in a typical histogram) used, thus, they are better at determining the distribution shape compared to Histogram. This is a brief tutorial on how to make density plot and some tips of how to make it prettier.

## Keep in Mind

Notice that the variable on the x-axis should be continuous. Density Plot has hard time to deal with discrete variable.

## Also Consider

You might also want to know how to make a histogram or a line graph, click [Histogram](https://lost-stats.github.io/Presentation/histograms.html) or [Line graph](https://lost-stats.github.io/Presentation/line_graphs.html) for more information.


# Implementations

## R

For this R demonstration, we are going to use `ggplot2` package to create density plot. Additionally, we will use the dataset `diamonds` that is natively available in R. 

To begin with this R demonstration, make sure that we install and load all the useful packages that we need it. We use the function `p_load()` in `pacman` package to help us to install and load the packages at one time. 

```{r}
#install and load necessary packages
if (!require("pacman")) install.packages("pacman")
pacman::p_load(ggplot2,viridis,RColorBrewer,tidyverse,ggthemes,ggpubr,datasets)
```

Next, in order to make a density plot, we are going to use `ggplot()`and `geom_density()` functions. For instance, we will specify `price` as our x-axis. 

```{r,warning=FALSE}

ggplot(diamonds, aes(x=price))+
  geom_density()
```


Then, we can always change the color of the density plot using the `col` argument and fill the color inside the density plot using `fill` argument. Furthermore, we can specify the degree of transparency density fill area using the argument `alpha` where `alpha` ranges from 0 to 1. 

```{r,warning=FALSE}
ggplot(diamonds, aes(x=price))+
  geom_density(fill="lightblue",
               col='black',
               alpha=0.6)
```


We can also change the type of line of the density plot as well by adding `linetype=` inside `geom_density()`.

```{r}
ggplot(diamonds, aes(x=price))+
  geom_density(fill="lightblue",
               col='black',
               linetype="dashed")
```


Furthermore, you can also combine both histogram and density plots together.

```{r,warning=FALSE,message=FALSE}
ggplot(diamonds, aes(x=price)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="grey45")+
 geom_density(col="red",size=1,linetype='dashed') 
```


What happen if we want to make multiple densities? 

For example, we want to make multiple densities plots for price based on the type of cut, all we need to do is adding `fill=cut` inside `aes()`.

```{r}
ggplot(data=diamonds, aes(x=price,fill=cut)) +
    geom_density(adjust=1.5, 
                 alpha=.3)
```

Congratulations!!!
It's time to start the adventure of the Density Plot!!!
:)
