---
title: "Propensity Score Matching"
author: "Jonas Bowman"
date: "5/31/2021"
output: md_document
parent: Matching
grand_parent: Model Estimation
has_children: false
nav_order: 1
mathjax: true ## Switch to false if this page has no equations or other math rendering.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Propensity Score Matching

Propensity Score Matching(PSM) is a non-parametric method of estimating a treatment effect in situations where randomization is not possible. This method comes from [Rosenbaum & Rubin, 1983](https://www.jstor.org/stable/2335942?seq=1) and works by estimating a propensity score which is the predicted probability that someone received treatment based on the explanatory variables of interest. This reduces bias in observational studies by balancing the covariates between a group that received the treatment and a group that did not, essentially attempting to replicate a randomized control trial.


## Basic Workflow

Step One:

run a logistic regression where the outcome variable is a binary indicator for whether or not someone received the treatment and gather the estimate of the propensity score. The explanatory variables in this case are the covariates that we might reasonably believe influence treatment

Step Two:

Match those that received treatment with those that did not based on propensity score. There are a number of different ways to perform this matching including, but not limited to :

*Nearest neighbor matching

*Exact matching

*Stratification matching

In this example we will focus on nearest neighor matching, which we will use through th MatchIt package

Step Three:

Check the balance of the matched sample using the summary command. This is good practice to ensure that the matching worked well and that the constructed control and treatment groups have similar observable characteristics.

Step Four;

Run analysis using this new sample which will give us an average treatment effect. 




## Keep in Mind

Propensity Score Matching is based on selection on observable characteristics. This assume that the potential outcome is independent of the treatment D conditional on the covariates. 
$Y_i(1),Y_i(0)\bot|X_i$

This is known as the Conditional Independence Assumption 

Propensity Score Matching also requires us to make the Common Support or Overlap Assumption.

$0<Pr(D_i = 1 | Xi = x)<1$

This assumption says that the probability that the treatment is equal to 1 for each level of x is between zero and one, or in other words there are both treated and untreated units for each level of x. 

Because Propensity Score Matching relies so heavily on these assumptions, [Regression Discontinuity](https://lost-stats.github.io/Model_Estimation/Research_Design/regression_discontinuity_design.html), [Differences in Differences](https://lost-stats.github.io/Model_Estimation/Research_Design/two_by_two_difference_in_difference.html), and other quasi-experimental designs may provide more accurate and robust results.




## Further Resouces

A great place to find more information about the MatchIt package is on the package's [github site](https://github.com/kosukeimai/MatchIt). 

Noah Greifer's [Cran Page](https://cran.r-project.org/web/packages/MatchIt/vignettes/MatchIt.html) is another good resource for getting started using Propensity Score Matchign in R. 





## Implemenation in R

Implementation can be done using the MatchIt package

Data comes from [OpenIntro.org](https://www.openintro.org/data/index.php?data=smoking)


```{r Implementing Propensity Score Matching, warning=FALSE, message=FALSE}
##load the packages we need.

library(pacman)
p_load(tidyverse, MatchIt)

##load data on smoking in the United Kingdom.

smoking = read_csv("smoking.csv")

##turn smoking into a binary variable.
smoking = smoking %>%mutate(smoke = if_else(smoke == "Yes",1,0))

##Step One: Run the logistic regression.

ps_model = glm( smoke ~ gender+age+marital_status+ethnicity+region, data=smoking)

#get a summary of the model
summary(ps_model)


## Step Two: Match on propensity score.

#does not apply in this situation, but need to make sure there are no missing values in the covariates we are choosing. 

#in order to match use the matchit command, passing the function a formula, the data to use and the method, in this case, nearest neighor estimation.

match = matchit(smoke ~ gender+age+marital_status+ethnicity+region, method = "nearest", data =smoking)

##Step 3: Check for Balance.
summary(match)

##create data frame from matches using the match.data function.

match_data = match.data(match)


#check the dimensions.
dim(match_data)


##Step Four: Conduct Analysis using the new sample.

##turn marital status into a factor variable so that we can use it in our regression 
match_data = match_data %>% mutate(marital_status = as.factor(marital_status))


## we can now get the treatment effect of smoking on gross income with and without controls
lm_nocontrols = lm(marital_status ~ smoke, data= match_data)

lm_nocontrols

# with controls
lm_controls =lm(marital_status ~ smoke+age+gender+ethnicity+marital_status, data=match_data)

lm_controls


```


